### 2019-07-06 thesis-template/dummy.tex
update

\documentclass{scrartcl}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{csquotes}
% Include Graphic-files:
\usepackage{graphicx}
\usepackage{caption}

\newcommand{\source}[1]{\vspace{-3pt} \caption*{ Source: {#1}} }

\begin{document}

\begin{figure}[!t]
\centering
  \begin{minipage}{0.4\textwidth}
  \includegraphics[width=0.9\textwidth]{img/heartDwnsmpl.png}
    \caption*{Glyphs}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.4\textwidth}
    \includegraphics[width=0.9\textwidth]{img/heartDwnsmpl-TFL.PNG}
     \caption*{Hlawatsch's approach: TFLs}
    \label{b)}
  \end{minipage}
\caption{\enquote{heart} test field}
\label{heart-ftle-comp}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Alternative: put content in separate files
% Check the difference between including these files using \input{filename} and \include{filename} and see which one you like better
%\chapter{Einleitung}\label{intro}
%\input{introduction}
%
%\chapter{Voraussetzungen}\label{bg}
%\input{background}

\end{document}